Experiences with Planning for Natural Language Generation

### 2008-07-25 trunk/Papers/experiences/domains.tex
changes to conclusions; added new footnote in domains section

The GIVE task comes with a number of interesting research challenges
both for the NLG and the planning communities.  On the NLG side, one
major problem is that the natural language instruction sequences can't
simply be obtained by verbalising each action instance in a plan in turn.
For instance, the NLG system could generate an instruction sequence
starting with ``turn left; walk forward; press the button; turn right; walk
forward; walk forward; turn right; walk forward; walk forward; turn left;
walk forward''.  However, such directions would be clumsy and boring, and
it is to be expected that frustrated users will quickly cancel an
interaction with such a system, resulting in a bad evaluation score.
Having just pressed the button, it would be much better to say ``turn
around and walk through the door''. That is, plans must be
\emph{summarised} by merging multiple action instances into a single
instruction, in order to present them to the user. Conversely, it may also
be necessary to \emph{elaborate} on a single planning step by expressing it
with several instructions.  Having just entered the top left room, it may
be easier for the user to understand the instruction ``walk to the centre
of the room; turn right; now press the green button in front of you'',
rather than the instruction ``press the green button on the wall to your
right''.  To plan the referring expression ``the green button in front of
you'' at a time when the button is not yet in front of the user, the NLG
system must keep track of the hypothetical changes in what objects are
visible to the user.  Thus the acts of referring and instructing, and the
modules for planning and NLG, must be tightly integrated.

>>>>

The GIVE task comes with a number of interesting research challenges
both for the NLG and the planning communities.  On the NLG side, one
major problem is that the natural language instruction sequences can't
simply be obtained by verbalising each action instance in a plan in turn.
For instance, the NLG system could generate an instruction sequence
starting with ``turn left; walk forward; press the button; turn right; walk
forward; walk forward; turn right; walk forward; walk forward; turn left;
walk forward''.  However, such directions would be clumsy and boring, and
it is to be expected that frustrated users will quickly cancel an
interaction with such a system, resulting in a bad evaluation score.
Having just pressed the button, it would be much better to say ``turn
around and walk through the door''. That is, plans must be
\emph{summarised} by merging multiple action instances into a single
instruction, in order to present them to the
user.\footnote{This task is similar to the construction of ``macro''
actions, but is primarily intended for plan presentation.}
Conversely, it may also be necessary to \emph{elaborate} on a single
planning step by expressing it with several instructions.  Having just
entered the top left room, it may be easier for the user to understand the
instruction ``walk to the centre of the room; turn right; now press the
green button in front of you'', rather than the instruction ``press the
green button on the wall to your right''.  To plan the referring expression
``the green button in front of you'' at a time when the button is not yet
in front of the user, the NLG system must keep track of the hypothetical
changes in what objects are visible to the user.  Thus the acts of
referring and instructing, and the modules for planning and NLG, must be
tightly integrated.
